---
title: "Self-assignment script - GT-seq 2021 manuscript - Kristen Hayward"
output: html_notebook
---

Tutorial:https://alexkychen.github.io/assignPOP/analyze.html

Install and load package:
```{r}
install.packages("assignPOP")
library(assignPOP)
```


For this manuscript, we evaluated self-assignment in two ways: (1) based on Management Units/MUs, and (2) based on STRUCTURE-determined genetic clusters. 


**1. SELF-ASSIGNMENT - MUs**

Don't forget to remove populations with <10 individuals prior to importing genetic data (GT-seq genotypes in genpop format). Import data: 
```{r}
genpop<- read.Genepop("file/path/genepop_file.txt", pop.names = c("pop1","pop2","pop3","etc."),haploid = FALSE)
```


Monte-Carlo cross-validation:
```{r}
assign.MC(genpop, train.inds = 0.75, train.loci = 0.75, iterations = 30, model = "svm", dir = "MU_results/")
```


Calculate assignment accuracy:
```{r}
accuMC<- accuracy.MC(dir = "MU_results/")
```


Can read results in R:
```{r}
accuMC<- read.table("MU_results/Rate_of_30_tests_11_pops.txt",header = T)
View(accuMC)
```


Visualize results:
```{r}
#create boxplot for assignment accuracy
library(ggplot2)
accuracy.plot(accuMC, pop = c("pop1","pop2","pop3","etc."))+
  ylim(0,1)+
  annotate("segment", x = 0.4, xend = 3.6, y = 0.33, yend = 0.33, colour = "firebrick", size = 1)+  #add red line at 0.3
  ggtitle("Monte-Carlo cross-validation using genetic loci")+ #add a plot title
  theme(plot.title = element_text(size = 20, face = "bold"))
```


Print mean and standard deviation across assignment tests:
```{r}
assign.matrix(dir = "file/path/to/folder/MU_results/")
```





**SELF-ASSIGNMENT - STRUCUTRE GENETIC CLUSTERS**

Import structure file:
```{r}
clusterfile<- read.Structure("file/path/to/structure/file/filename.txt")

#keep in mind which pop is which, for us: 
#pop1 = arctic archipelago
#pop2 = polar basin
#pop3 = hudson complex
#pop4 = m'clintok channel
```


Monte-Carlo cross-validation:
```{r}
assign.MC(clusterfile, train.inds = 0.75, train.loci = 0.75, iterations = 30, model = "svm", dir = "cluster_results/")
```


Calculate assignment accuracy:
```{r}
accuMC_clusters<- accuracy.MC(dir = "cluster_results/")
```


Can read results in R:
```{r}
accuMC_clusters<- read.table("cluster_results/Rate_of_30_tests_4_pops.txt",header = T)
View(accuMC_clusters)
```


Visualize results:
```{r}
#create boxplot for assignment accuracy
library(ggplot2)
accuracy.plot(accuMC_clusters, pop = c(1,2,3,4))+
  ylim(0,1)+
  annotate("segment", x = 0.4, xend = 3.6, y = 0.33, yend = 0.33, colour = "firebrick", size = 1)+  #add red line at 0.3
  ggtitle("Monte-Carlo cross-validation using genetic loci")+ #add a plot title
  theme(plot.title = element_text(size = 20, face = "bold"))
```


Print mean and standard deviation across assignment tests:
```{r}
assign.matrix(dir = "file/path/to/cluster_results/")
```

